---
title: "Fig2_Rversion"
author: "Bre Rivera Waterman"
date: "2024-02-22"
output: github_document
editor_options: 
  chunk_output_type: console
chunk_output_type: inline
---

```{r setup, include=FALSE}
install.packages("knitr", repos='http://cran.us.r-project.org', quiet=TRUE)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

current_path <- rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path))
setwd("./../")
getwd()
```

 - import libraries

```{r}
library(tidyverse)
library(colourpicker)
library(lubridate)
library(emmeans)
library(sf)
library(ggspatial)
library(spData)
library(ggpmisc)
#install.packages("ggpubr")
library("rstudioapi")
#install.packages("ggExtra")
library(ggExtra)
library(patchwork)


```



## Scatter plot  
- using new merged dataset to create scatter plot
- add kernel densities to the plots (along the side)

```{r}
#some filtering/mutating first 
#C:/Users/b923w423/Documents/GitHub/ICON-ModEx_Open_Manuscript


combods_split <- read_csv("ICON-ModEx_Combined_Predicted_Observed_Respiration_Rates.csv", na = c(-9999, "N/A", "NA", NA)) %>%
#remova Na's
  #split data set into "hot" and "cold" spots
  mutate(ds_split = ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7, "500+",                         ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7, "500-", "500-") )) %>% 
  
#clasifying by quadrant (hot, cold, misclass hot, misclass cold) by year
  mutate(ds_quadrant23 =                                                    ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate_Nov2023 > 2.7 , "hs23",
  ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate_Nov2023 < 2.7 , "misclass_hs23",                                                                                              
ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate_Nov2023 < 2.7 , "cs23",    
ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate_Nov2023 > 2.7 , "misclass_cs23", NA)))) ,                   
      
ds_quadrant19 =        
 ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate_Sep2019 > 2.7 , "hs19",
  ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate_Sep2019 < 2.7 , "misclass_hs19",  
 ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate_Sep2019 > 2.7 , "misclass_cs19",
ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate_Sep2019 < 2.7 , "cs19", NA ) )))) 

 
```


### Scatter plot
 - needs to have kernel density plots on y and x axis
 - needs use log data
 - needs to have overall line, line for above 500, below 500, and R^2 values for all three
 - facet wrap or color dots by iteration (first versus last)
 
 - right now s19s iteration, but want OG 0th iteration

```{r}
#this will change - waiting on real 0th iteration data from Stefan/Brie
firstiterationds <- combods_filt %>% filter(str_detect(Sample_Name, "S19S"))

#run 3 lm up here, then extract R^2, then call into annotate 
lm_all <- lm(Log_Predicted_Normalized_Respiration_Rate_Sep2019 ~ Log_Observed_Normalized_Respiration_Rate, firstiterationds)
summary(lm_all)

lm_sup <- lm(Log_Predicted_Normalized_Respiration_Rate_Sep2019 ~ Log_Observed_Normalized_Respiration_Rate, firstiterationds %>% filter(ds_split == "500+"))
summary(lm_sup)

lm_sub <- lm(Log_Predicted_Normalized_Respiration_Rate_Sep2019 ~ Log_Observed_Normalized_Respiration_Rate, firstiterationds %>% filter(ds_split == "500-"))
summary(lm_sub)

#plot
fullscatter_firstiteration <- 
 firstiterationds %>%
    ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate ),
                      abs(Log_Predicted_Normalized_Respiration_Rate_Sep2019) )) +
                      #,  color =  ds_split))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point(color = "darkgrey", size = 3, pch = 21) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
 
   
        geom_abline (linetype = "longdash", color="black") + #not a clear threshold 
   stat_poly_line(color = "black", alpha = 0.5) + #overall best fit
 # stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
    geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate), #above/below threshold best fit
                      abs(Log_Predicted_Normalized_Respiration_Rate_Sep2019),
                      color =  ds_split), method = "lm", alpha = 0.2) + 
  xlab("log10(Observed rates)") + 
  ylab("log10(Predicted rates)") + 
scale_shape_manual(values = c(21, 16)) + #shape or color by iteration
  scale_color_manual(values = c( "lightblue", "maroon")) +

annotate('text', label = str_c('R^2 == ', round(summary(lm_all)$r.squared, 2)), parse = T, x = 3.5, y = 0.75, size = 5, family = 'serif', color = "black")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sub)$r.squared, 2)), parse = T, x = 3.5, y = 0.5, size = 5, family = 'serif', color="lightblue")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sup)$r.squared, 2)), parse = T, x = 3.5, y = 0.25, size = 5, family = 'serif',  color="maroon")+
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16,  family = 'serif'), 
         legend.position = 0)
fullscatter_firstiteration

#adding density plots on axes
ggExtra::ggMarginal(fullscatter_firstiteration, fill = "lightgrey")

#save file
ggsave(filename = "Scripts/fullscatter_firstiteration.tiff", plot = fullscatter_firstiteration, width = 6, height = 6, dpi = 300)   
 
 
```



 - final iteration 
 
```{r}
#add in lm's
#run 3 lm up here, then extract R^2, then call into annotate 
lm_all_final <- lm(Log_Predicted_Normalized_Respiration_Rate_Nov2023 ~ Log_Observed_Normalized_Respiration_Rate, combods_split)
summary(lm_all_final)

lm_sup_final <- lm(Log_Predicted_Normalized_Respiration_Rate_Nov2023 ~ Log_Observed_Normalized_Respiration_Rate, combods_split %>% filter(ds_split == "500+"))
summary(lm_sup_final)

lm_sub_final <- lm(Log_Predicted_Normalized_Respiration_Rate_Nov2023 ~ Log_Observed_Normalized_Respiration_Rate, combods_split %>% filter(ds_split == "500-"))
summary(lm_sub_final)



fullscatter_lastiteration <- 
  combods_split %>% ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate ),
                      abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023) #)) + #
                      ,  color =  ds_split))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point(color = "darkgrey", size = 3, pch = 21) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
      geom_abline (linetype = "longdash", color="black") + #not a clear threshold 
   stat_poly_line(color = "black", alpha = 0.5) + #overall best fit
 # stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
 
    geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
                      abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023),
                      color =  ds_split), method = "lm", alpha = 0.2) + 
  xlab("log10(Observed rates)") + 
  ylab("log10(Predicted rates)") + 
scale_shape_manual(values = c(21, 16)) + #shape or color by iteration
  scale_color_manual(values = c( "lightblue", "maroon")) +

annotate('text', label = str_c('R^2 == ', round(summary(lm_all_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.75, size = 5, family = 'serif', color = "black")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sub_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.5, size = 5, family = 'serif', color="lightblue")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sup_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.25, size = 5, family = 'serif',  color="maroon")+
    
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16, family = 'serif'), 
         legend.position = 0)

#adding density plots on axes
 ggExtra::ggMarginal(fullscatter_lastiteration, fill = "lightgrey")
 

ggsave(filename = "Scripts/fullscatter_lastiteration.tiff", plot = fullscatter_lastiteration, width = 6, height = 6, dpi = 300)   
```




- exploring misclass patterns
```{r}
combods_date <- combods_split %>%
   mutate(Date_whole = ymd(Observed_Sample_Date), Year = year(Observed_Sample_Date), Month = month(Observed_Sample_Date), Day = day(Observed_Sample_Date))
 


#based on previous plots, temporal outliers are from 2023 only
combods_date %>% ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate),
                      abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023), color = Year ))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point( pch = 21, size = 3) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
      geom_abline (linetype = "longdash", color="black") + #not a clear threshold 
  # stat_poly_line(color = "black", alpha = 0.5) + #overall best fit
  #stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
  # scale_color_gradientn(colours = rev(rainbow(12))) +

   # geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
    #                  abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023),
     #                 color =  ds_split), method = "lm", alpha = 0.2) +
  
  xlab("log10(Observed rates)") + 
  ylab("log10(Predicted rates)") + 
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16))


  
  
```


- checking classifications
```{r}
#to check that classifications are correct
combods_date %>% ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate),
                      abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023),  color =  ds_quadrant23))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point( pch = 21) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
      geom_abline (linetype = "longdash", color="black") + #not a clear threshold 
  # stat_poly_line(color = "black", alpha = 0.5) + #overall best fit
  #stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
   #scale_color_gradientn(colours = rev(rainbow(12))) +

   # geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
    #                  abs(Log_Predicted_Normalized_Respiration_Rate_Nov2023),
     #                 color =  ds_split), method = "lm", alpha = 0.2) +
  
  xlab("log10(Observed rates+1)") + 
  ylab("log10(Predicted rates+1)") + 
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16))

  
 
combods_date %>% ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate),
                      abs(Log_Predicted_Normalized_Respiration_Rate_Sep2019),  color = ds_quadrant19 ))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point( pch = 21) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
      geom_abline (linetype = "longdash", color="black") + #not a clear threshold 
  # stat_poly_line(color = "black", alpha = 0.5) + #overall best fit
  #stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
  # scale_color_gradientn(colours = rev(rainbow(12))) +

   # geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
                     # abs(Log_Predicted_Normalized_Respiration_Rate),
                     # color =  ds_split), method = "lm", alpha = 0.2) +
  
  xlab("log10(Observed rates+1)") + 
  ylab("log10(Predicted rates+1)") + 
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16))
```


#map to see miss-classifications spatially
  - NA's not filtered out?
```{r}
usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE))

outliers2023 <- 
ggplot(usa) +
  geom_sf(color = "#2b2b2b", fill = "lightgrey", size=0.125, alpha = 0.15) +
  coord_sf(crs = st_crs("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"), datum = NA) +
   geom_spatial_point(data=combods_date, aes(x=Longitude, y=Latitude, color = ds_quadrant23), size = 3) + 
               scale_color_manual(values=c(   "#B9D3EE", "maroon", "darkgrey", "#CDBA96" ) )+
     ylab("") + 
  xlab("Longitude") + 
    theme_minimal() +
   labs(title="2023 iteration")
outliers2023

outliers2019 <- 
ggplot(usa) +
  geom_sf(color = "#2b2b2b", fill = "lightgrey", size=0.125, alpha = 0.15) +
  coord_sf(crs = st_crs("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"), datum = NA) +
   geom_spatial_point(data=combods_date, aes(x=Longitude, y=Latitude, color = ds_quadrant19), size = 3) + 
               scale_color_manual(values=c(   "#B9D3EE", "maroon", "darkgrey", "#CDBA96" ) )+
     ylab("Latitude") + 
  xlab("Longitude") + 
    theme_minimal() + 
  theme(legend.position = 0)+
     labs(title="2022 iteration")



misclassmap <- outliers2019 + outliers2023
misclassmap


ggsave(filename = "Scripts/misclassmap.tiff", plot = misclassmap, width = 10, height = 5, dpi = 300)   

```

