---
title: An R Markdown document converted from "make_scatter_plot.ipynb"
output: html_document
---


```{r}
# Parameters cell for papermill launching
work_dir <- "./intermediate_branch_data/Nov-2023-log10"
```

```{r, setup}
# Change to working directory
setwd(work_dir)

# Confirmation of change directory
getwd()
```

```{r}
library(tidyverse)
library(ggextra)
```

## Scatter plot  
- using new merged dataset to create scatter plot
- add kernel densities to the plots (along the side)

```{r}
#some filtering/mutating first 
combods_split <- read_csv("ICON-ModEx_Combined_Predicted_Observed_Respiration_Rates.csv", na = c(-9999, "N/A", "NA", NA)) %>%
#remova Na's
  #split data set into "hot" and "cold" spots
  mutate(ds_split = ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7, "500+",                         ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7, "500-", "500-") )) %>% 
  
#clasifying by quadrant (hot, cold, misclass hot, misclass cold) by year
# Need to get rid of the year because we are working year-by-year
  mutate(ds_quadrant23 =                                                    ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate > 2.7 , "hs23",
  ifelse(Log_Observed_Normalized_Respiration_Rate > 2.7 & Log_Predicted_Normalized_Respiration_Rate < 2.7 , "misclass_hs23",                                                                                              
ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate < 2.7 , "cs23",    
ifelse(Log_Observed_Normalized_Respiration_Rate < 2.7 & Log_Predicted_Normalized_Respiration_Rate > 2.7 , "misclass_cs23", NA)))) )
```

### Scatter plot
 - needs to have kernel density plots on y and x axis
 - needs use log data
 - needs to have overall line, line for above 500, below 500, and R^2 values for all three
 - facet wrap or color dots by iteration (first versus last)
 
 - right now s19s iteration, but want OG 0th iteration

```{r}
#add in lm's
#run 3 lm up here, then extract R^2, then call into annotate 
lm_all_final <- lm(Log_Predicted_Normalized_Respiration_Rate ~ Log_Observed_Normalized_Respiration_Rate, combods_split)
summary(lm_all_final)

lm_sup_final <- lm(Log_Predicted_Normalized_Respiration_Rate ~ Log_Observed_Normalized_Respiration_Rate, combods_split %>% filter(ds_split == "500+"))
summary(lm_sup_final)

lm_sub_final <- lm(Log_Predicted_Normalized_Respiration_Rate ~ Log_Observed_Normalized_Respiration_Rate, combods_split %>% filter(ds_split == "500-"))
summary(lm_sub_final)



fullscatter_lastiteration <- 
  combods_split %>% ggplot(aes(abs(Log_Observed_Normalized_Respiration_Rate ),
                      abs(Log_Predicted_Normalized_Respiration_Rate) #)) + #
                      ,  color =  ds_split))+ 
#to see R^2 for above/below thresholds, remove color grouping above and use stat poly eq fx (below)
  geom_point(color = "darkgrey", size = 3, pch = 21) + 
    geom_vline (xintercept = 2.7, linetype = "dashed", color="grey") + 
    geom_hline (yintercept = 2.7, linetype = "dashed", color="dark grey") + #not a clear threshold 
      geom_abline (linetype = "longdash", color="black") + #not a clear threshold

# Required by ggextra, fails to import above. Where are the other lines (blue and red) plotted?
     #ggpmisc::stat_poly_line(color = "black", alpha = 0.5) + #overall best fit


# stat_poly_eq(use_label(c( "R2"))) + #remove hashtag to see R^2 of log500 split
 
    geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
                      abs(Log_Predicted_Normalized_Respiration_Rate),
                      color =  ds_split), method = "lm", alpha = 0.2) + 
    geom_smooth( aes(abs(Log_Observed_Normalized_Respiration_Rate ), #above/below threshold best fit
                      abs(Log_Predicted_Normalized_Respiration_Rate)), method = "lm", alpha = 0.2,
                      color =  "black") +
  xlab("log10(Observed rates)") + 
  ylab("log10(Predicted rates)") + 
scale_shape_manual(values = c(21, 16)) + #shape or color by iteration
  scale_color_manual(values = c( "lightblue", "maroon")) +

annotate('text', label = str_c('R^2 == ', round(summary(lm_all_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.75, size = 5, family = 'serif', color = "black")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sub_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.5, size = 5, family = 'serif', color="lightblue")+
annotate('text', label = str_c('R^2 == ', round(summary(lm_sup_final)$r.squared, 2)), parse = T, x = 3.5, y = 0.25, size = 5, family = 'serif',  color="maroon")+
    
  lims(y = c(0,4), x = c(0,4)) + 
  theme_classic() +
   theme(text = element_text(size = 16, family = 'serif'), 
         legend.position = 0)

#adding density plots on axes
 histograms <- ggExtra::ggMarginal(fullscatter_lastiteration, fill = "lightgrey")
 
# Original plot save in tidy
ggsave(filename = "./scatter.png", plot = histograms, width = 6, height = 6, dpi = 300)  
#ggsave(filename = "./scatter.pdf", plot = fullscatter_lastiteration, width = 6, height = 6, units="in", dpi = 300, device="pdf")  
# View plot inline
#histograms

# Alternative in base
#png("./scatter.png")
#print(histograms)
#print(fullscatter_lastiteration)
#dev.off()
```



```{r}
# Display figure inline if necessary
histograms
```

