---
title: "ICON_Mapping"
author: "Bre Rivera Waterman"
date: "2023-09-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping R file


 - import libraries

```{r}
library(tidyverse)
library(colourpicker)
library(lubridate)
library(emmeans)
library(sf)
library(ggspatial)
library(spData)
library(rnaturalearth)
library(ggpmisc)
#install.packages("ggpubr")
library(rnaturalearthdata)
library("rstudioapi")

current_path <- rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path))
setwd("./../")
getwd()

```


- Loading metadata files & binding rows 
```{r}
metadata1 <- read.csv("./v8_WHONDRS_S19S_Sediment/WHONDRS_S19S_Sediment_Metadata/v4_WHONDRS_S19S_Metadata.csv") %>% slice(-1) %>% 
  #C:/Users/b923w423/Documents/GitHub/ICON-ModEx_Open_Manuscript/v8_WHONDRS_S19S_Sediment/WHONDRS_S19S_Sediment_Metadata/v4_WHONDRS_S19S_Metadata.csv
  
  #C:/Users/b923w423/Documents/GitHub/ICON-ModEx_Open_Manuscript/v8_WHONDRS_S19S_Sediment/WHONDRS_S19S_Sediment_Metadata/v4_WHONDRS_S19S_Metadata.csv
                                                                dplyr::select(lat = "MS_Latitude_dec.deg",
                                                                long = "MS_Longitude_dec.deg",
                                                                "Date",
                                                                site_ID = "Sample_ID",
                                                                site_name = "Stream_Name",
                                                                Agency = "Associated_Site",
                                                                state = "State_or_Province", 
                                                                Country = "Country") %>%
                                                                 mutate(site_ac = "no")
metadata2 <- read.csv("C:/Users/b923w423/Documents/GitHub/ICON-ModEx_Open_Manuscript/v3_CM_SSS_Data_Package/v3_CM_SSS_Field_Metadata.csv", header = TRUE) %>%      mutate(Country = "USA")  %>%                                 
                                                                dplyr::select(lat = "Sample_Latitude",
                                                                long = "Sample_Longitude",
                                                                Date = "Sample_Date",
                                                                site_ID = "Parent_ID",
                                                                site_name = "Stream_Name",
                                                                Agency = "Organization",
                                                                state = "State",
                                                                 "Country",
                                                                site_ac = "Site_ID")
                                                               
                                                            


combo_metadata <- rbind(metadata1,metadata2) %>%  #combining datasets
   filter(Country == "USA", state != "Puerto Rico", state !="Alaska") %>% #next few lines are to determine which sites are NEON, create a time series Y or N column based on if sites are NEON
   mutate(Date_whole = ymd(Date), Year = year(Date),
          lat = as.numeric(lat), long = as.numeric(long),
  Time_series = case_when(
    str_detect(site_ac, "HOPB|MAYF|MART") ~"Yes", 
    TRUE ~ "No"),
#create season col based on months
   season = case_when(
   month(Date) %in% c(3,4,5) ~ "Spring", 
   month(Date) %in% c(6,7,8) ~ "Summer", 
   month(Date) %in% c(9,10,11) ~ "Fall", 
   TRUE ~ "Winter", 
))


```



## Fig 1; seasonal map with time series site different shape
```{r}

usa <- st_as_sf(maps::map("state", fill=TRUE, plot =FALSE))

ggplot(usa) +
  geom_sf(color = "#2b2b2b", fill = "#FCFCFC", size=0.125) +
  coord_sf(crs = st_crs("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs")) +  #geom_spatial_point(data=combo_metadata, aes(x=long,y=lat, fill = season),  pch = 21, size = 4) +  
   geom_spatial_point(data=combo_metadata, aes(x=long,y=lat, fill = season, shape = Time_series, size = Time_series)) + 

  scale_fill_manual(values =  c("#8c510a","#dfc27d", "#c7eae5", "#01665e" ))+ #color blind safe colors
  #scale_color_manual(values =  c("#8B2252", "#EE799F", "#4d9221", "#e6f5d0" ))+ #color blind safe colors
  scale_shape_manual(values = c(21, 15)) + 
  scale_size_manual(values = c(5, 8)) +
  ylab("Latitude") + 
  xlab("Longitude") + 
   annotation_scale(location = "bl", width_hint = 0.3,  pad_x = unit(0.25, "in"), pad_y = unit(0.25, "in")) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.5, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
  ggtitle("Sites sampled across CONUS") +
  theme_bw() + 
      theme(panel.grid.major = element_line(color = gray(0.9), linetype = "dashed", 
        size = 0.1), panel.background = element_rect(fill = "white"), text = element_text(size = 18)) +
  labs(fill = "Season", shape = "Time series") + 
  guides(size = "none", fill = guide_legend(override.aes = list(size = 5, shape = 21)),  shape = guide_legend(override.aes = list(size = 5)))  
c("#E0EEEE")


```
 
  - needs scale bar
  - pretty basic map, gets the point across but could be improved and open to suggestions!
  - interactive map tutorial linked here: http://maxpohlman.com/rgistutorial


## Scatter plot  
- using new merged dataset to create scatter plot
```{r}
combods <- read.csv("C:/Users/b923w423/Documents/GitHub/ICON-ModEx_Open_Manuscript/ICON-ModEx_Combined_Predicted_Observed_Respiration_Rates.csv") %>% filter(!is.na(Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment),!is.na(Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment)) %>% filter(Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment > -9990, Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment > -9990 )

#scatter plot
fullscatter <- 
  combods %>% ggplot(aes(abs(Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment), abs(Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment))) +
  geom_point(size = 4, alpha = 0.5) + 
    geom_abline (slope=1, linetype = "dashed", color="Red") + 
      geom_text(x = 500, y = 500, label = "*", size = 15, color="orange") + 
   stat_poly_line(color = "light grey", alpha = 0.01) +
  stat_poly_eq(use_label(c( "R2"))) +
  xlab("Observed rates") + 
  ylab("Predicted rates") + 
    lims(y = c(0,4000), x = c(0,4000)) + 
        geom_text(x = 500, y = 500, label = "*", size = 15, color="orange") + 
  theme_classic() +
   theme(text = element_text(size = 18))


under500 <- combods %>%  
  ggplot(aes(abs(Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment), abs(Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment))) +
  geom_point(size = 4, alpha = 0.5) + 
    geom_abline (slope=1, linetype = "dashed", color="Red") + 
   stat_poly_line(color = "light grey", alpha = 0.01) +
  stat_poly_eq(use_label(c( "R2")), label.x = "right", label.y = "bottom") +
  ylab("") + 
  xlab("Observed rates") + 
    lims(y = c(0,500), x = c(0,500)) + 
  theme_classic() + 
     theme(text = element_text(size = 18))


over500 <- combods %>% 
 ggplot(aes(abs(Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment), abs(Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment))) +
  geom_point(size = 4, alpha = 0.5) + 
    geom_abline (slope=1, linetype = "dashed", color="Red") + 
    stat_poly_line(color = "light grey", alpha = 0.01) +
  stat_poly_eq(use_label(c( "R2")), label.x = "right", label.y = "bottom") +
  ylab("") + 
  xlab("") + 
    lims(y = c(500,4000), x = c(500,4000)) + 
  theme_classic() + 
     theme(text = element_text(size = 18))

fullscatter + (over500 / under500)


```
 
  - add units to axis?
  - increase text size



## Fig 4 
```{r}
combods <- combods %>% filter(Latitude < 50 & Latitude > 20) %>% mutate(LogObs = Observed_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment, LogPred = Predicted_Normalized_Respiration_Rate_mg_DO_per_H_per_L_sediment)

library(RColorBrewer)


Obs <- ggplot(usa) +
  geom_sf(color = "#2b2b2b", fill = "#FCFCFC", size=0.125) +
  coord_sf(crs = st_crs("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"), datum = NA) +
  geom_spatial_point(data=combods, aes(x=Longitude,y=Latitude, fill = log(-LogObs+1)), pch = 21, size =4) +     
  #scale_shape_manual(values = c(16, 8)) +
   # scale_size_manual(values = c(4, 8)) +
  scale_fill_distiller(palette = "Spectral",  #YlOrRd
                        limits = c(min(log(0+1)), max(log(4000+10)) )) + 
  ylab("") + 
  xlab("") +
   ggtitle("Observed respiration rates at sampled sites") +
 theme_classic() +
   theme(panel.background = element_rect(fill = "white"), text = element_text(size = 12), legend.position ="none") +
    labs(fill = "log(Resp. + 5)") 

#aliceblue
  
 Pred<- 
  ggplot(usa) +
  geom_sf(color = "#2b2b2b", fill = "#FCFCFC", size=0.125) +
  coord_sf(crs = st_crs("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs"), datum = NA) +
  geom_spatial_point(data=combods, aes(x=Longitude,y=Latitude, fill = log(-LogPred+1)), pch = 21, size =4) +     
  #scale_shape_manual(values = c(16, 8)) +
   # scale_size_manual(values = c(4, 8)) +
  scale_fill_distiller( palette = "Spectral", 
                        limits = c(min(log(0+1)), max(log(4000+10)) ) )  + 
  ylab("") + 
  xlab("")  +
   ggtitle("Predicted respiration rates at sampled sites") +
     #   annotation_scale(location = "bl", width_hint = 0.3,  pad_x = unit(0.1, "in"), pad_y = unit(0.1, "in")) +
  theme_classic() +
   theme(panel.background = element_rect(fill = "white"), text = element_text(size = 12), legend.position = "bottom") +
       labs(fill = "log(Resp. + 1)") 


  Obs / Pred

```


